---
title: "ingest_clean"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(data.table)
library(magrittr)
```

To-do list, CDC data:

* Copy the variable names and column widths for (?) 2013, 2014, 2016 into variables here; import all CDC data.
* Select variables of interest
* For each variable, decode CDC's shorthand into the data format and datatypes we desire
* Also decode "flags" as needed
* Join all years

To-do list, NBER data:

* Read.csv?
* Select features of interest
* Decode any shorthand(?) as needed
* Join all years

To-do list, all:

* Standardized our feature/variable names, and MERGE them all
* (Not crucial) feature and correlation analysis on the final dataset

# Ingesting data from the CDC and NBER

## CDC Data

All CDC data was retrieved through their Vital Statistics online portal: https://www.cdc.gov/nchs/data_access/vitalstatsonline.htm#Fetal_Death. 

Thank goodness

```{r}
codebook_base <- tribble(
  ~width, ~name,
  6, "F",
  1, "Version",
  2, "F",
  1, "Obstetric Estimate Tabulation Flag",
  4, "Delivery Year",
  2, "Delivery Month",
  4, "F",
  4, "Delivery Time",
  1, "Weekday",
  2, "Territory of Occurence",
  2, "F",
  3, "County FIPS",
  1, "County Population",
  1, "Delivery Place (revised)",
  1, "Delivery Place",
  16, "F",
  1, "Delivery Place Recode",
  31, "F",
  1, "Mother Age Imputed",
  1, "Reported Age of Mother Flag",
  2, "Mother Single Year of Age",
  2, "Mother Age Recode 14",
  1, "Mother Age Recode 9",
  2, "Mother Birth Country",
  4, "F",
  1, "Mother Birth State Recode",
  4, "F",
  2, "Mother Residence Postal State",
  3, "Mother FIPS County",
  5, "F",
  1, "Population of Residence County",
  17, "F",
  1, "Record Type",
  1, "Residence Status",
  2, "Mother Race Recode 31",
  1, "Mother Race Recode 6",
  2, "Mother Race Recode 15",
  1, "Mother Bridged Race Recode",
  1, "Mother Race Imputed",
  3, "F",
  1, "Mother Hispanic Origin",
  1, "Mother Bridged Race/Hispanic Origin",
  1, "Mother Race/Hispanic Origin",
  1, "Mother Education Revised",
  26, "F",
  1, "Father Reported Age Used",
  4, "F",
  2, "Father Combined Age",
  2, "Father Age Recode 11",
  2, "Prior Births Now Living",
  2, "Prior Births Now Dead",
  2, "F",
  1, "Live Birth Order Recode",
  9, "F",
  3, "Interval Since Last Live Birth Recode",
  2, "Interval Since Last Live Birth Recode 11",
  2, "Month Prenatal Care Began (Revised)",
  1, "Month Prenatal Care Began Recode (Revised)",
  24, "F",
  1, "WIC",
  2, "Cigarettes Before Pregnancy",
  2, "Cigarettes 1st Trimester",
  2, "Cigarettes 2nd Trimester",
  2, "Cigarettes 3rd Trimester",
  1, "Cigarette Recode (Revised)",
  4, "F",
  2, "Mother Height (in)",
  4, "BMI",
  1, "BMI Recode",
  3, "F",
  3, "Prepregnancy Weight Recode",
  1, "F",
  1, "Prepregnancy Diabetes",
  1, "Gestational Diabetes",
  1, "Prepregnancy Hpertension",
  1, "Gestational Hypertension",
  1, "Hypertension Eclampsia",
  1, "Infertility Treatment",
  1, "Fertility Enhancing Drugs",
  1, "Asst Reproductive Technology",
  1, "Previous Cesareans",
  2, "Previous Cesareans Num",
  1, "F",
  1, "Diabetes",
  1, "Chronic Hypertension",
  1, "Pregnancy Associated Hypertension",
  1, "Eclampsia",
  1, "Breech Delivery",
  1, "Fetal Presentation",
  1, "Route and Method of Delivery",
  1, "Trial of Labour Attempted",
  1, "Delivery Method Recode (Revised)",
  1, "Forceps",
  1, "Vacuum",
  1, "Delivery Method Recode",
  1, "Ruptured Uterus",
  1, "Admit to Intensive Care",
  1, "Attendant",
  17, "F",
  1, "Plurality Recode",
  1, "F",
  1, "Plurality Imputed",
  2, "F",
  1, "Set Order Recode",
  9, "F",
  1, "Sex of Infant",
  1, "Sex Imputed",
  2, "Last Normal Menses Month",
  2, "F",
  4, "Last Normal Menses Year",
  3, "F",
  1, "Gestation Imputed Flag",
  1, "Obstetric Estimate of Gestation Using Flag",
  2, "Gestation Detail in Weeks",
  2, "Gestation Recode 12",
  1, "Gestation Recode 5",
  2, "Obstetric Gestation",
  1, "F",
  1, "Combined Gestation Used Flag",
  2, "Obstetric Estimate Edited",
  2, "Obstetric Estimate Recode 12",
  1, "Obstetric Estimate Recode 5",
  4, "F",
  4, "Birth Weight Grams",
  2, "Birth Weight Recode 14",
  1, "Birth Weight Recode 4",
  1, "F",
  1, "Estimated Time of Fetal Death",
  1, "Autopsy Performed",
  1, "Histological Placental Exam Performed",
  1, "Autopsy/Hist Results used in Cause",
  11, "F",
  21, "Flags1",
  10, "F",
  15, "Flags2",
  6, "F",
  4, "Flags3",
  1, "F",
  1, "Flags4",
  1, "F",
  3, "Flags5",
  1, "F",
  6, "Flags6",
  2610, "F"
)

codebook_base2013 <- tribble(
  ~width, ~name,
  6, "F",
  1, "Version",
  1, "F",
  1, "Tabulation Flag",
  5, "F",
  4, "Delivery Year",
  2, "Delivery Month",
  8, "F",
  1, "Weekday",
  2, "Territory of Occurence",
  5, "F",
  3, "County FIPS",
  1, "County Population",
  1, "F",
  1, "Delivery Place",
  16, "F",
  1, "Delivery Place Recode",
  27, "F",
  1, "Mother Age Imputed",
  1, "Reported Age of Mother Flag",
  2, "Mother Single Year of Age",
  2, "Mother Age Recode 14",
  1, "Mother Age Recode 9",
  15, "F",
  2, "Mother's Territory of Residence",
  3, "F",
  3, "Mother FIPS County",
  15, "F",
  1, "Population of Residence County",
  4, "F",
  1, "Record Type",
  1, "Residence Status",
  4, "F",
  1, "Mother Race Recode",
  1, "Mother Race Imputed",
  3, "F",
  1, "Mother Hispanic Origin",
  1, "Mother Race/Hispanic Origin",
  3, "F",
  1, "Mother Marital Status",
  21, "F",
  1, "Father Reported Age Used",
  10, "F",
  2, "Father Age Recode 11",
  24, "F",
  1, "Live Birth Order Recode",
  4, "F",
  1, "Total Birth Order Recode",
  58, "F",
  2, "Weight Gain",
  1, "Weight Gain Recode",
  52, "F",
  1, "Diabetes",
  3, "F",
  1, "Chronic Hypertension",
  1, "Prepregnancy Hpertension",
  1, "Eclampsia",
  46, "F",
  1, "Breech Delivery",
  15, "F",
  1, "Forceps",
  1, "Vacuum",
  1, "Hyesterectomy",
  1, "Delivery Method Recode",
  6, "F",
  1, "Attendant",
  12, "F",
  1, "Plurality",
  1, "F",
  1, "Plurality Imputed",
  10, "F",
  1, "Sex",
  1, "Sex Imputed",
  2, "F",
  2, "Last Normal Menses Month",
  4, "Last Normal Menses Year",
  2, "Obstetric Gestation",
  3, "F",
  2, "Gestation Detail in Weeks",
  2, "Gestation Recode 12",
  1, "Gestation Recode 5",
  1, "Obstetric Estimate of Gestation Using Flag",
  1, "Gestation Imputed Flag",
  5, "F",
  4, "Birth Weight Grams",
  4, "F",
  2, "Birth Weight Recode 14",
  1, "Birth Weight Recode 4",
  30, "F",
  1, "Anencephalus",
  1, "Spina Bifida",
  7, "F",
  1, "Omphalocele",
  4, "F",
  1, "Cleft Lip",
  4, "F",
  1, "Downs Syndrome",
  45, "F",
  232, "Flags")

# not used right now
end_flags <-
  c("Education of Mother",
    "Obstetric Estimate of Gestation",
    "Tobaccu Use",
    "Mother Height",
    "Pregnancy Weight",
    "WIC",
    "Prepregnancy diabetes",
    "Gestational diabetes",
    "prepregnancy hypertension",
    "gestational hypertension",
    "hypertension eclampsia",
    "infertility treatment",
    "fertility enhance drug",
    "reproductive tech",
    "prev cesarean",
    "num of prev cesarean",
    "fetal presentation",
    "final route and method of delivery",
    "trial of labour attempted",
    "ruptured uterus",
    "admission to ICU",
    "month prenatal care began",
    "combined gestation",
    "cig before preg",
    "cig before 1",
    "cig before 2",
    "cig before 3",
    "facility",
    "delivery method recode",
    "diabetes flag",
    "chron hyper flag",
    "preg assoc hyper flag",
    "eclampsia flag",
    "breech flag",
    "forceps flag",
    "vacuum flag",
    "mother hispanic origin flag",
    "mother race revised flag",
    "father age flag",
    "live birth month flag",
    "F",
    "last normal menses month flag",
    "F",
    "last normal menses year flag",
    "mother age flag",
    "attendant flag",
    "F",
    "estimated time of fetal death (revised)",
    "autopsy flag",
    "histological placement flag",
    "mothers bridged race recode flag",
    "num of prev births flag")
```


### 2013

Accompanying codebook: ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/DVS/fetaldeath/2013FetalUserGuide.pdf

```{r}
fd_2013 <- read.fwf(
   "../data/cdc/VS13FetalDetailPubUS.txt",
   widths = codebook_base2013$width,
   col.names = codebook_base2013$name,
   header = FALSE,
   stringsAsFactors = FALSE,
   n = 100
 )
```

### 2014

Accompanying codebook: ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/DVS/fetaldeath/2014FetalUserGuide.pdf

```{r}
fd_2014 <- read.fwf(
  "data/cdc/VS14FetalPubUS_NEW.txt",
  widths = codebook_base$width,
  col.names = codebook_base$name,
  header = FALSE,
  stringsAsFactors = FALSE,
  n = 100
)
```


### 2015

Accompanying codebook: ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/DVS/fetaldeath/2015FetalUserGuide.pdf

```{r}
fd_2015 <- read.fwf(
  "data/cdc/Fetal2015US.txt", 
  widths = codebook_base$width,
  col.names = codebook_base$name,
  header = FALSE,
  stringsAsFactors = FALSE,
  n = 100
)
```

### 2016

Accompanying codebook: ftp://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/DVS/fetaldeath/2016FetalUserGuide.pdf

```{r}
fd_2016 <- read.fwf(
  "data/cdc/2016_FET_Analysis_Public_US_20190409.txt", 
  widths = codebook_base$width,
  col.names = codebook_base$name,
  header = FALSE,
  stringsAsFactors = FALSE,
  n = 100
)
```


## NBER Data

This is also really CDC data from the same source, but already parsed into a much-nicer csv file by the Natl Bureau of Economic Research: https://www.nber.org/research/data/vital-statistics-natality-birth-data

## 

```{r}

renamed_columns <- c(
  "Version",
  "DeliveryYear",
  "Age",
  "Race",
  "Education",
  "PrevTerminations",
  "WIC",
  "PrepregnancySmoking",
  "BMI",
  "Height",
  "Weight",
  "Parity",
  "PrepregnancyDiabetes",
  "GestationalDiabetes",
 "PrepregnancyHypertension",
  "GestationalHypertension",
  "HypertensionEclampsia",
  "PrevPretermBirths",
  "InfertilityTreatment",
  "InfertilityDrugs",
  "ART",
  "PrevCesareans",
  "Gonorrhea",
  "Syphilis",
  "Chlamydia",
  "HepatitisB",
  "HepatitisC",
  "Stillborn",
  "Preterm"
)


natl2013 <- fread("../data/cdc/natl2013.csv") 


natl2013_clean <- natl2013 %>% 
  mutate(Stillborn = 0) %>% 
  select(revision, dob_yy, mager, mracerec, meduc, priorterm, wic, cig_0, bmi, m_ht_in, pwgt_r, lbo_rec, rf_diab, rf_gest, rf_phyp, rf_ghyp, rf_eclam, rf_ppterm, rf_inftr, rf_fedrg, rf_artec, rf_cesar, ip_gono, ip_syph, ip_chlam, ip_hepatb, ip_hepatc, Stillborn, gestrec3) %>%
  set_colnames(renamed_columns)

fwrite(natl2013_clean, "../data/clean/natl2013_clean.csv")

remove(list = ls()) #so that you don't saturate your RAM

natl2014 <- fread("../data/cdc/natl2014.csv") 

natl2014_clean <- natl2014 %>% 
  mutate(Stillborn = 0) %>%
  mutate(revision = 0) %>%
  select(revision, dob_yy, mager, mbrace, meduc, priorterm, wic, cig_0, bmi, m_ht_in, pwgt_r, lbo_rec, rf_diab, rf_gest, rf_phyp, rf_ghyp, rf_eclam, rf_ppterm, rf_inftr, rf_fedrg, rf_artec, rf_cesar, ip_gono, ip_syph, ip_chlam, ip_hepatb, ip_hepatc, Stillborn, gestrec3) %>%
  set_colnames(renamed_columns)


```