---
title: "fairchild_fake_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's make some fake data

```{r cars}
library(fakeR)
library(tidyverse)
library(recipes)
```

```{r pressure, echo=FALSE}
preg_df <- data.frame(age=c(15,26,21,23),
                      race=c("white", "black", "asian", "amerindian"),
                      married=c(0,1,1,0),
                      education=c("High School", "College","Doctorate","8thGrade"),
                      prev_term=c(1,1,2,0),
                      WIC=c(0,1,0,1),
                      smoking=c(0,1,7,20),
                      BMI=c(18.5,21.2,18.1,27.8),
                      height=c(62,51,72,64),
                      weight=c(150,122,143,111),
                      parity=c(1,1,1,0),
                      diabetes=c(1,0,1,0),
                      gdiabetes=c(0,1,1,0),
                      hypertension=c(1,0,1,0),
                      eclampsia=c(0,0,0,1),
                      pptb=c(0,0,0,1),
                      infertility=c(1,0,0,0),
                      idrugs=c(0,1,0,0),
                      cesarean=c(0,0,1,0),
                      gonorrhea=c(1,1,0,0),
                      chlamydia=c(0,0,0,1),
                      hepb=c(0,1,0,0),
                      hepc=c(1,1,0,0),
                      preterm=c(0,0,28,25),
                      stillborn=c(0,1,0,0))
preg_sim <- simulate_dataset(preg_df, n = 16000)


```

## Preprocessing Step

1) Applying inclusion criteria, each obs must be complete.
2) Mothers must be 18 or older, material morbidity excluded
3) Gestational Age in alive babies less than 21 weeks excluded
4) Multiple birth pregnancies excluded
5) Pregnancies that ended in fetal death due to external causes excluded.

6) Parity status of the mother was deducted from the number of prior births variable
7) A new class variable was annotated to specify the outcome of pregnancies more accurately. 
8) Fetal death cases were divided into late and early stillbirth based on their gestational age, at 28 weeks.
9) Early stillbirth cases of less than 21 weeks were excluded because they are clinically defined as miscarriage cases. 
10) Live births were divided into uncomplicated pregnancies, and PTB pregnancies at 37 weeks.
11) We should get like 12,000,000 normal pregnancies, 1,000,000 PTB cases, 7924 early stillbirth cases and 8310 late stillbirth cases.

12) Continuous Variable mean-zero normalization and unit-variance normalization. 
13) Nominal predictors were one-hot encoded
14) CDC data was partitioned into four sets; feature selection data, training data, validation data and test data along 10, 70, 10, 10 split.


```{r}
#functions
normalize <- function(x){
  return ((x-mean(x))/sd(x))
}
```

```{r}
preg_sim <- preg_sim %>% drop_na() %>% #1
  filter(age > 18) %>% #2
  filter(preterm > 21) %>% #3
#4 #5 NOT DONE
  mutate(BMI = normalize(BMI)) %>% #12
  mutate(height = normalize(height)) %>%
  mutate(weight = normalize(weight)) %>%
  mutate(id = row_number()) %>% #adding an id
  mutate(outcomes = ifelse(stillborn == 1,
                           ifelse(preterm < 21, "miscarriage",
                                  ifelse(preterm < 28, "early_stillbirth", "late_stillbirth")),
                           ifelse(preterm < 37, "preterm", "normal"))) %>% #7-11
  mutate(race = paste("race", race, sep = "_"), #13
         race_hot = 1,
         education = paste("education", education, sep = "_"),
         education_hot = 1) %>%
  spread(key = race, value = race_hot, fill = 0) %>%
  spread(key = education, value = education_hot, fill = 0)
  

```
